---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Raul"
date: "April 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

logit <- function(x, plot = TRUE) {
    
    if(!is.numeric(x)) { print("please provide a numeric value or vector") }
    else {
      y <- log(1/(1-x)) #<-apply logistic transformation
      if(plot == TRUE) {qqplot(y,min(x):max(x))} #<-optional visual aid
      return(y)
    }
}

inv.logit <- function(x, plot = TRUE) {
    
    if(!is.numeric(x)) { print("please provide a numeric value or vector") }
    else {
      y <- exp(x)/(1 + exp(x)) #<-apply inverse logistic transformation
      if(plot == TRUE) {qqplot(y,min(x):max(x))} #<-optional visual aid
      return(y)
    }
}

```


# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   


# Question:

Is this a good model?  (Write your answer here.)


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

