---
title: 'NYCFlights: Arrival Delay Logictic Model'
author: "Raul"
date: "April 29, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

logit <- function(x, plot = TRUE) {
    
    if(!is.numeric(x)) { print("please provide a numeric value or vector") }
    else {
      y <- log(1/(1-x)) #<-apply logistic transformation
      if(plot == TRUE) {qqplot(min(x):max(x),y)} #<-optional visual aid
      return(y)
    }
}

inv.logit <- function(x, plot = TRUE) {
    
    if(!is.numeric(x)) { print("please provide a numeric value or vector") }
    else {
      y <- exp(x)/(1 + exp(x)) #<-apply inverse logistic transformation
      if(plot == TRUE) {qqplot(min(x):max(x),y)} #<-optional visual aid
      return(y)
    }
}

x <- -10:10
logit(x)
inv.logit(x)

```


# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   

```{r}
library(dplyr)
library(readr)

# load the data
data <- read.csv("C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/03-linear-regression/03-exercise-nycflights-model/NYCflights.csv", stringsAsFactors = FALSE)

# We want to model arrival delay as a binary variable. 
# In other words, we consider a flight to be delayed if its arrival is 22 minutes delayed (y=1), otherwise it's considered on time (y=0).
data$y <- as.numeric(data$arr_delay >= 22)

# Tabulation of our dependent variable
table(data$y)
```

Borrowing from last weeks' assigment, we concluded that the best liner model that minimized MSE--by the AIC measure--was:

arr_delay ~ dep_delay + month + air_time + distance + carrier + lat.dest + lon.dest

We now use that model, but changing the dependent variable with our newly created binary dependent variable

``` {r warning = FALSE}
OLS <- lm(y ~ dep_delay + month.flight + air_time + distance + carrier + lat.dest + lon.dest, data = data)

summary(OLS)

# obtain predictions
pred.OLS <- predict(OLS)
summary(pred.OLS)

# We see that OLS predicted values outside [0,1]. We now turn to our logit model:
LOGIT <- glm(y ~ dep_delay + month.flight + air_time + distance + carrier + lat.dest + lon.dest, data = data, family = binomial(link = "logit"))
summary(LOGIT)
pred.LOGIT <- predict(LOGIT, type = 'response')

```

# Question:

Is this a good model? 

One way to assess how good a model is to determine 'goodness of fit'

```{r waning = FALSE}
# To answer how good the model is, I interpret several measures of fit:
# Deviance, pseudoR2, ANOVA of independent variable coefficients
pscl::pR2(LOGIT)
anova(LOGIT, test = 'Chisq')

# ANOVA test the difference between the null deviance (naive model)
# against the deviance of our model. Residual deviance keeps going
# down as we add variables to the model. 
# The drops in deviance are statistically significant.
```

We cannot assess a model without seeing how 'accurate' it is.
To assess accuracy, we turn to confusion matrix and ROC curves.

``` {r warning = FALSE}
# Also, I select the 0.5 as a threshold to determine when a y=1 and when y=0.
# Based on that, I can then measure how accurate my model is. 
pred.LOGIT <- ifelse(pred.LOGIT > 0.5, 1, 0)
misClasificError <- mean(pred.LOGIT != LOGIT$y)
print(paste('Accuracy', 1 - misClasificError))

library(caret)
confusionMatrix(pred.LOGIT, LOGIT$y, positive = '1')

```


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

