---
title: "Improving Model Perfromance / Tuning Parameters"
author: "Raul Torres"
date: "`r Sys.Date()`"
output: html_document
---


## Tuning Parameter

Generically and regardless of model type, what are the purposes of a model
tuning parameters?

The purpuse is to run the model several times, each time adjusting one or more 
(depending on the number of parameters a model could have), in order to see which parameter (or set of parameters)
improves the model fit statistic of our chioce. 

For example, for a knn model-which can take on parameter 'k', we would 'tune' this parameter by selecting several values
of k and seeing which one maximizes the kappa statistic. 



## Caret Models

This assignment demonstrates the use of caret for constructing models. Each
model should be built and compared using using `Kappa` as the performance
metric calculated using 10-fold repeated cross-validation with 3 folds.

Using the rectangular data that you created for the NYCFlights to create a model
for arr_delay >= 15 minutes

- glm
- rpart
- knn
- C50
- randomForest
- adaBoost
- Two methods of your choice from the Caret Model List (you will need to install any dependencies)

Save the caret objects with the names provided.

## Loading libraries and data
```{r}
library(caret)

# Import dataset
data <- read.csv("C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/03-linear-regression/03-exercise-nycflights-model/NYCflights.csv", stringsAsFactors = FALSE)

# Create a binary dependent variable that is 1 when flight is late (arr_delay > 22 minutes), 0 otherwise. 
data$y <- as.numeric(data$arr_delay >= 15)

# Remove rows where predictions won't be fesible due to missingness in covariates (or response variable)
data <- na.omit(data[,c("y","dep_delay","month.flight","air_time","distance","carrier","lat.dest","lon.dest","arr_delay")])

```

## Tune C5.0 model
```{r}

# first, we select control options
ctr <- trainControl(method = "cv", number = 10, selectionFunction = "best")

# next, we select our two models along with values for their respective parameters and store them in a grid
grid_c50 <- expand.grid(.model = "C50",
                        .trials = c(10, 15, 20),
                        .winnow = "FALSE")

fit.c50 <- train(y ~ ., data = data, method = "C5.0", metric = "Kappa", trControl = ctrl, tuneGrid = grid_c50)

```


## Tune Random Forest model
```{r}

# next, we select our two models along with values for their respective parameters and store them in a grid
grid_rf <- expand.grid(.mtry = c(2, 4))

fit.rf <- train(y ~ ., data = data, method = "rf", metric = "Kappa", trControl = ctrl, tuneGrid = grid_rf)

```


## Compare the  models?
Note that we use Kappa as the fit measure on which to declare a model 'best'. Also, we are not going for the
most parsimonous model, for the one that maximizes Kappa. 

```{r}
fit.c50
fit.rf
```

## Which is best?  Why?
Based on the model (and set of parameters within that model) that maximized Kappa, the best model is:



