---
title: "Classification Metrics"
author: "Raul Torres"
date: "May 9, 2017"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions"}

rates <- function(y, yhat) {
    tp <- sum(yhat==1 & y==1) / length(y==1)
    fp <- sum(yhat==1 & y==0) / length(y==0)
    tn <- sum(yhat==0 & y==0) / length(y==0)
    fn <- sum(yhat==0 & y==1) / length(y==1)
    list(TP.rate = tp, FP.rate = fp, TN.rate = tn, FN.rate = fn)
}


```

## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}

data <- read.csv("C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/03-linear-regression/03-exercise-nycflights-model/NYCflights.csv", stringsAsFactors = FALSE)
```

As a reminder this logit models late arrival (an arrival of 22 or more minutes after scheduled arrival) controling for several covariates

```{r}
data$y <- as.numeric(data$arr_delay >= 22)
LOGIT <- glm(y ~ dep_delay + month.flight + air_time + distance + carrier + lat.dest + lon.dest, 
             data = data, 
             family = binomial(link = "logit"))
summary(LOGIT)

```

## Obtaining the vector of predicted values setting a threshold of 0.5
``` {r}
yprob <- predict(LOGIT, type = 'response')
yhat <- as.numeric(yprob>=0.5)
y <- LOGIT$y
```

## Obtaining the four requested rates: 
```{r}
rates(y, yhat)
```

