---
title: "Classification Metrics"
author: "Raul Torres"
date: "May 9, 2017"
=======
title: "Classification Mettrics"
author: "Christopher Brown"
date: "October 5, 2015"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions"}

rates <- function(y, yhat) {
   tp <- sum(yhat==1 & y==1) / sum(y==1)
   fp <- sum(yhat==1 & y==0) / sum(y==0)
   tn <- sum(yhat==0 & y==0) / sum(y==0)
   fn <- sum(yhat==0 & y==1) / sum(y==1)
   sensi <- sum(y==1 & yhat==1) / (sum(y==1 & yhat==1) + sum(y==1 & yhat==0))
   speci <- sum(y==0 & yhat==0) / (sum(y==0 & yhat==0) + sum(y==0 & yhat==1))
   recall <- sensi
   precis <- sum(y==1 & yhat==1) / (sum(y==1 & yhat==1) + sum(y==0 & yhat==1))
   prev <- sum(y==1) / length(y)
   accur <- sum(yhat == y) / length(y)
   E <-  ((prod(sum(y == 1),sum(yhat == 1)) / length(y)) +
          (prod(sum(y == 0),sum(yhat == 0)) / length(y))) / length(y)
   kappa <- (accur - E) / (1 - E)

   list(TruePos.rate = tp,
        FalsePos.rate = fp,
        TrueNeg.rate = tn,
        FalseNeg.rate = fn,
        sensitivity = sensi,
        specificity = speci,
        recall = recall,
        precision = precis,
        prevalance = prev,
        accuracy = accur,
        kappa = kappa)
}

```{r "definitions", echo=FALSE}
```

=======
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}
data <- read.csv("C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/03-linear-regression/03-exercise-nycflights-model/NYCflights.csv", stringsAsFactors = FALSE)
```

As a reminder this logit models late arrival (an arrival of 22 or more minutes after scheduled arrival) controling for several covariates

```{r}
data$y <- as.numeric(data$arr_delay >= 22)
LOGIT <- glm(y ~ dep_delay + month.flight + air_time + distance + carrier + lat.dest + lon.dest, 
             data = data, 
             family = binomial(link = "logit"))

```

## Obtaining the vector of predicted values setting a threshold of 0.5
``` {r}
yprob <- predict(LOGIT, type = 'response')
yhat <- as.numeric(yprob>=0.5)
y <- LOGIT$y
```

## Obtaining the four requested rates: 
```{r}
rates(y, yhat)
```

