---
title: "Assingment NYC Flight Data"
author: "Raul Torres"
date: "April 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
library("readr")

# reading all CSV datasets and saving them as df objects in memory
data_path <- "C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/"
df_airports <- read.csv(paste0(data_path,"02-building-blocks/02-exercise-nycflights/data/airports.csv"), stringsAsFactors=FALSE)
df_flights <- read.csv(paste0(data_path,"02-building-blocks/02-exercise-nycflights/data/flights.csv"), stringsAsFactors=FALSE)
df_planes <- read.csv(paste0(data_path,"02-building-blocks/02-exercise-nycflights/data/planes.csv"), stringsAsFactors=FALSE)
df_weather <- read.csv(paste0(data_path,"02-building-blocks/02-exercise-nycflights/data/weather.csv"), stringsAsFactors=FALSE)

```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
hist(df_flights$arr_delay, xlab = "arrival delays", main = "NYC airport arrival delays")
hist(df_flights$dep_delay, xlab = "departure delays", main = "NYC airport departure delays")
```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

# first we create a date variable based on existing variables month, day, and year
df_flights$date <- as.Date(
                       paste0(
                           as.character(df_flights$month),"-",as.character(df_flights$day),"-",as.character(df_flights$year)
                       ),
                       "%m-%d-%Y"
                   )

# plot newly created date variable and arrival delay
plot(df_flights$date, df_flights$arr_delay, xlab = "Date", ylab = "Arrival delay")
```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}
library(dplyr)
df_flightsNoNA <- df_flights[is.na(df_flights$flight) == FALSE,] #remove missing values in variable flights
mytable <-
 df_flightsNoNA %>% 
  group_by(origin, carrier) %>% #<- select by-grp
   mutate(totals = sum(flight)) %>% #<- create variable on the fly containing the sum of by-grp
    select(one_of(c("origin", "carrier", "totals"))) %>% #<- drop unwanted variables
      arrange(origin, carrier) %>% #<-sort by origin and carrier
       summarise(count = first(totals)) #<- collapse dataframe leaving the first value of by-grp
mytable

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
# Joining data frames with respect to 'fligts'
# This means we'll retain all the rows in flights (and matched rows in other dataframes),
# but we'll drop rows unique to other dataframes. 
# In dplyr lingo, we'll left_join with flights being the 'left' dataframe

# create a common key between flights & airports
df_airports$origin <- df_airports$faa #<-creating a key variable for merging with flights

#df_flights <- sample_frac(df_flights, 0.25, replace = TRUE) used for testing - commented out for now

# create a common key between flights & weather
df_weather$date <- as.Date(
                       paste0(
                           as.character(df_weather$month),"-",as.character(df_weather$day),"-",as.character(df_weather$year)
                       ),
                       "%m-%d-%Y"
                   )

df_merged <- 
 left_join(df_flights, df_airports, by = "origin") %>%
   left_join(df_planes, by = "tailnum") %>%
     left_join(df_weather, by = "date")

identical(dim(df_flights)[1],dim(df_merged)[1])
print(paste0("We added ", dim(df_merged)[2] - dim(df_flights)[2], " new variables as we joined dataframes."))


```



