---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "Raul Torres"
date: "April 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: Load data and plot dependent variable 

```{r "Step 1: Load data and plot dependent variable" }
    
p <- "C:/Users/rtorres/Desktop/MachLearn/MyUser/CSX460/03-linear-regression/03-exercise-nycflights-model"  
df_orig <- read.csv(paste0(p,"/NYCflights.csv"))
df <- df_orig
hist(df$arr_delay)
df$date <- as.character(df$date)
df$xmas <- df$date %in% c("2013-12-21","2013-12-22","2013-12-23","2013-12-24","2013-12-26")

# remove cases where dependent variable is missing
df <- df[is.na(df$arr_delay)==FALSE,]

# remove variables that don't vary
truevars <- sapply(df, function(x) length(unique(x)) > 1)
df <- df[, truevars]

# remove variables that are only NA and one factor
for (i in 1:dim(df)[2]) {
    
    if (length(unique(df[,i])) <= 2) {
     u <- unique(df[,i])
     p <- paste0("variable:", names(df)[i]," has these unique values: ")
     print(p)
     print(u)
     df[,i] <- NULL
    }
    
}

dim(df_orig)
dim(df)


# remove 

# in order to have a more Normal distribution of "y", we transform to inverse of y
df$arr_delay_inv <- (df$arr_delay^(-1))
hist(df$arr_delay_inv)
# because changing to inverse would make the impacts on y more difficult to interpret, we stick to y.

```


## Step 2.a: Fit the naive model as baseline. Display results.
```{r "Step 2: Fit the dummy model and select another model. Display results" }

model.dummy <- lm(arr_delay ~ 1, df) #<-same as ybar
mean(df$arr_delay, na.rm = TRUE)
summary(model.dummy)
res.dummy <- resid(model.dummy)
predicted <- fitted(model.dummy)
summary(res.dummy)
```


Thoughts about naive model:


## Step 2.b: Fit the saturated model. Display results.
```{r}

model.sat <- lm(arr_delay ~ ., data = df)
summary(model.sat)
res.sat <- resid(model.sat)
predicted <- fitted(model.sat)
```
# Post estimation plots
```{r}
plot(df[is.na(df$arr_delay)==FALSE,"arr_delay"], predicted, xlab = "observed values", ylab = "predicted values")
plot(predicted, res.sat, xlab = "predicted values", ylab = "residuals")
```

Thoughts about saturated model:


## Step 3: Remove collinear features
```{r}
library(caret)
X <- select(df, -arr_delay)
correlations <- cor(X)
correlations[1:6, 1:6]

highcorr <- findCorrelation(correlations, cutoff = .70, names = TRUE)
highcorr

```


## Step 4: Fit a new (hopefully better) model
```{r "Step 4: Fit a new (hopefully better) model" }



model2 <- lm(arr_delay ~ carrier + distance + month.flight + precip + wind_gust + wind_speed + speed + dest + origin + xmas, data = df)
summary(model2)

res.model2 <- resid(model2)
hist(res.model2)

```

# Question:

Is this a good model?  
 

